---
title: "cm007 Exercises: Practice with `dplyr`"
output: 
  html_document:
    keep_md: true
    theme: paper
---

<!---The following chunk allows errors when knitting--->

```{r allow errors, echo = FALSE}
knitr::opts_chunk$set(error = TRUE, warning = FALSE)
```

```{r}
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(gapminder))
suppressPackageStartupMessages(library(tsibble))
```


This worksheet contains exercises aimed for practice with `dplyr`. 
```{r summarize demo}
gapminder %>% 
  summarize(mu    = mean(lifeExp),
            sigma = sd(lifeExp))
```


```{r group-by demo}
gapminder %>% 
  group_by(continent, year)
```

```{r group_by and summarize}
gapminder %>% 
  group_by(continent, year) %>% 
  summarize(mu    = mean(lifeExp),
            sigma = sd(lifeExp))
```

```{r grouped summarize}
gapminder %>% 
  group_by(country) %>% 
  summarize(n = n())
```


```{r HOMEWORK TIP!}
gap_inc <- gapminder %>% 
  arrange(year) %>% 
  group_by(country) %>%
  mutate(gdpPercap_inc = gdpPercap - lag(gdpPercap))
DT::datatable(gap_inc)
```

gap_inc %>% 
  tidyr::drop_na()
```{r removes 'NA'}
gap_inc %>% 
  tidyr::drop_na()
```

1. (a) What's the minimum life expectancy for each continent and each year? (b) Add the corresponding country to the tibble, too. (c) Arrange by min life expectancy.

```{r}
gapminder %>% 
  group_by(continent, year) %>% 
    summarize(min_life = min(lifeExp),
            country = country [lifeExp == min_life]) %>%
    arrange(min_life)

```

2. Calculate the growth in population since the first year on record _for each country_ by rearranging the following lines, and filling in the `FILL_THIS_IN`. Here's another convenience function for you: `dplyr::first()`. 


```{r}
gapminder %>%
  group_by(country) %>%
  arrange(year) %>%
  mutate(rel_growth = pop - first(pop)) %>%
  DT::datatable()
```


3. Determine the country that experienced the sharpest 5-year drop in life expectancy, in each continent, sorted by the drop, by rearranging the following lines of code. Ensure there are no `NA`'s. Instead of using `lag()`, use the convenience function provided by the `tsibble` package, `tsibble::difference()`:

```


```

```{r complex dplyr demo}
gapminder %>% 
  group_by(country) %>% 
  arrange(year) %>% 
  mutate(inc_life_exp = difference(lifeExp)) %>%
  drop_na() %>% 
  ungroup() %>% 
  group_by(continent) %>% 
  filter(inc_life_exp == min(inc_life_exp)) %>% 
  arrange(inc_life_exp) %>% 
  knitr::kable()
  
```


